---
title: "Final Project/Research"
output: html_document
---

```{r, include= FALSE}
library(tidyverse)
library(janitor)
library(broom)
library(GGally)
library(purrr)
library(readxl)
library(dplyr)
library(vegan)
library(modelr)
```

### Currently, I am studying the recruitment characteristics of Arctomecon humulis; an endangered species endemic to Washington County Utah.  The dwarf bear poppy grows on gypsum outcrops covered in cryptobiotic crusts.  I have collected data on roughly 200 seedlings from three populations at the square foot plot and emergence point levels.  

```{r include=FALSE, message=FALSE, warning=FALSE}
beehive<-read_excel("./beehive_clean.xlsx")
wd<- read_excel("./wd_clean.xlsx")
rb<- read_excel("./rb_clean.xlsx")
wb<- full_join(beehive, wd)
full<- full_join(wb,rb)

beehive_plots <- beehive %>% distinct(tag_number, .keep_all = TRUE)
wd_plots <- wd %>% distinct(tag_number, .keep_all = TRUE)
rb_plots <- rb %>% distinct(tag_number, .keep_all = TRUE)
wb_plots <- full_join(beehive_plots, wd_plots)
full_plots <- full_join(wb_plots, rb_plots)
```

  The plots below help illustrate some of the findings at the plot level.
 
```{r, echo= FALSE}
ggplot(full_plots, aes(x = disturbance, fill= population)) + 
  geom_bar() +
  theme_minimal() +
  labs(x= "Disturbance", y= "Number of Disturbed Plots") +
  theme(axis.text.x = element_text(angle=90)) +
  theme(axis.text.x = element_text(size= 6)) +
  facet_wrap(~population)
```

   The disturbance regimes of each population and the presence of disturbance in each of these sample plots varies as well.  
  
```{r, echo=FALSE}
 ggplot(full_plots, aes(x = topography_1, fill= population)) + 
  geom_bar() +
  theme_minimal() +
  labs(x= "Topography 1", y= "Number of Plots") +
  theme(axis.text.x = element_text(angle=90)) +
  theme(axis.text.x = element_text(size= 6)) +
  facet_wrap(~population)
```
 
   In this plot, flat refers to only a few degrees of slope along the plot, gradual is anywhere between about 10ยบ and 45ยบ, and steep refers to plots on slopes greater than 45ยบ. 
  
```{r, echo=FALSE}
ggplot(full_plots, aes(x = topography_2, fill= population)) + 
  geom_bar() +
  theme_minimal() +
  labs(x= "Topography 2", y= "Number of Plots") +
  theme(axis.text.x = element_text(angle=90)) +
  theme(axis.text.x = element_text(size= 5 )) +
  facet_wrap(~population)
```
 
   In this plot, frontslopes refer to either north or east facing slopes while backslopes refer to either south or west facing slopes. 
  
```{r, echo=FALSE}
ggplot(full_plots, aes(x = roughness, fill= population)) + 
  geom_bar() +
  theme_minimal() +
  labs(x= "Roughness", y= "Number of Plots") +
  theme(axis.text.x = element_text(angle=90)) +
  theme(axis.text.x = element_text(size= 6)) +
  facet_wrap(~population)
```

  The roughness values represent vertical "something" within the plot.  1 is 1cm or less difference.  2 is between 1cm and 3cm. 3 is 3cm or greater.  
 

  The cover of these plots varied and were different between populations.
 
```{r, echo=FALSE, warning=FALSE, message= FALSE, include=FALSE}
cover1 <- full_plots %>%select(population,tag_number,ends_with("_c"))

cover1[is.na(cover1)] <- 0

cover_matrix1 <- cover1 %>% 
  select(ends_with("_c"))

row.names(cover1) <- cover1$plant_id

NMDS2 <- metaMDS(cover_matrix1,
                distance = "jaccard",
                binary=TRUE,
                autotransform = FALSE)
MDS21 <- NMDS2$points[,1]
MDS22 <- NMDS2$points[,2]

```

```{r, echo=FALSE}
ggplot(cover1, aes(x=MDS21,y=MDS22,color=population)) +
  geom_point(size=4,alpha=.2) +
  stat_ellipse()

perm2 <- adonis(formula = cover_matrix1 ~ cover1$population)
perm2

```


  Significance of Cryptobiotic crust cover as a function of population.
```{r,}
crypto_c_mod <- aov(data=cover1,
                    formula = crypto_c ~ population)
summary(crypto_c_mod)

```

```{r, echo=FALSE}
TukeyHSD(crypto_c_mod) %>% plot()
```


 Significance of small gravel (less than a cm in diameter) cover as a function of population.
```{r,}
sm_gr_c_mod <- aov(data=cover1,
                   formula = sm_gr_c ~ population)
summary(sm_gr_c_mod)


```

```{r, echo=FALSE}
TukeyHSD(sm_gr_c_mod) %>% plot()
```



  Significance of large gravel (greater than a cm in diameter) cover as a function of population.
```{r, }
lg_gr_c_mod <- aov(data=cover1,
                   formula = lg_gr_c ~ population)
summary(lg_gr_c_mod)



```

```{r, echo=FALSE}
TukeyHSD(lg_gr_c_mod) %>% plot()
```

  Significance of exposed gypsum cover as a function of population.
```{r,}
gyp_c_mod <- aov(data=cover1,
                 formula = gyp_c ~ population)
summary(gyp_c_mod)



```

```{r, echo=FALSE}
TukeyHSD(gyp_c_mod) %>% plot()
```


   Significance of neighboring plant cover as a function of population.
```{r,}
plant_c_mod <- aov(data=cover1,
                 formula = plant_c ~ population)
summary(plant_c_mod)


```
 
```{r, echo=FALSE}
TukeyHSD(plant_c_mod) %>% plot()
```
 
   Significance of bare soil cover as a function of population.
```{r,}
bare_c_mod <- aov(data=cover1,
                 formula = bare_c ~ population)
summary(bare_c_mod)
```

```{r, echo=FALSE}
TukeyHSD(bare_c_mod) %>% plot()
```


### Each of these cover materials was significantly different among populations.  

___

### At the emergence point level, plants recruited into and grew from different materials within and between populations.

```{r, echo=FALSE, warning=FALSE, include=FALSE}
emergence1 <-full %>% 
  select(population,plant_number,ends_with("_e")) %>% 
  select(!contains("plant_e")) %>% 
  select(!contains("rivulet")) %>% 
  select(!contains("other")) %>% 
  mutate(plant_id = paste0(population,"_",plant_number)) 

emergence1[is.na(emergence1)] <- 0

emergence_matrix1 <- emergence1 %>% 
  select(ends_with("_e"))

row.names(emergence1) <- emergence1$plant_id


NMDS <- metaMDS(emergence_matrix1,
                distance = "jaccard",
                binary=TRUE,
                autotransform = FALSE)
MDS1 <- NMDS$points[,1]
MDS2 <- NMDS$points[,2]

emergence1$MDS1 <- MDS1
emergence1$MDS2 <- MDS2
names(emergence1)

```

```{r,echo=FALSE}
ggplot(emergence1, aes(x=MDS1,y=MDS2,color=population)) +
  geom_point(size=4,alpha=.2) +
  stat_ellipse()

```


  Significance of cryptobiotic soil as a emergence substrate as a function of population.
```{r, echo=FALSE}
emergence11 <- emergence1 %>% mutate_all(as.logical)
emergence_matrix2 <-emergence_matrix1 %>% mutate_all(as.logical)
emergence_matrix2$cow_e <- emergence_matrix2$plant_e <- emergence_matrix2$other_e <- emergence_matrix2$rivulet_e <- NULL

perm <- glm(formula= emergence11$plant_number~ crypto_e + lg_gr_e + sm_gr_e + gyp_e + bare_e, 
            family = binomial, data= emergence_matrix2, maxit= 100)
add_predictions(emergence_matrix2, perm, type= "response")

```

 